Digital economy in the UK; a multi-scalar story of the diffusion of web technologies

### 2023-03-22 paper/help/old/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Digital economy in the UK: a multi-scalar story of the diffusion
of web technologies}
\author[1]{Emmanouil Tranos%
\corref{cor1}%
}
 \ead{e.tranos@bristol.ac.uk}

\affiliation[1]{organization={University of Bristol and The Alan Turing
Institute},postcode={UK},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper maps the participation in the digital economy and its
evolution in the UK over space and time. Most of the existing economic
geography literature which dealt with the spatiality of the internet
employed supply-side measures, such as infrastructural capacity, in
order to understand the geography of the digital economy and its
potential spatial economic effects. Useful as these approaches might
have been, they cannot capture the micro-processes and the
characteristics of the individual online behaviour. Using large volumes
of archived and geolocated web content, this paper models the diffusion
of web technologies over space and time in the UK. The data and
geolocation strategy allow to capture these processes at a very granular
spatial scale. The modelling approach, which is based on simple spatial
analytical methods and on the estimation of diffusion curves at various
scales, enables to depict the role of geography and other cognitive
factors which drove the diffusion of web technologies. Although the
focus is on a recent historical period -- 1996-2012 -- the results of
the analysis depict diffusion mechanisms which can be very useful in
understanding the evolutionary patterns of the adoption of other newer
technologies.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{sec:1}{%
\section{Introduction}\label{sec:1}}

Geographers were always interested in how new technologies and
innovations diffuse across space and, importantly, how such
spatio-temporal processes can be modelled. The seminal contribution of
\citet{hagerstrand1968innovation} is illustrative of this early
interest. However, the torch of exploring and modelling such processes
had been passed to other disciplines such as economics, business studies
and sociology well before the `cultural turn' of economic geography
\citep{perkins2005international}. A potential explanation of lack of
studies exploring the diffusion of new and, more specifically, digital
technologies across \emph{both} space and time can be attributed to the
scarcity of relevant and granular enough data. As
\citet{zook2022mapping} highlight, digital activities are hardly ever
captured in official data.

@@

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time use this information to make predictions
regarding the diffusion of related future technologies. As per
\citet{meade2021modelling} a variety of actors have a direct interest in
gaining such knowledge including network equipment suppliers; network
operators, regulatory and local authorities. And despite the spatial
heterogeneity of such diffusion mechanisms and the policy relevance,
there are very limited attempts in the literature to analyse the
diffusion of new digital technologies at a granular level.

@@

\citet{zook2022mapping} Because new digital activities are rarely---if
ever---captured in official state data, researchers must rely on
information gathered from alternative sources. With this in mind, the
overarching aim of this article is outlining an approach for analyzing
the ``new spaces and geographies'' of digital phenomena and practices.

@@

\citet{fritsch2015new} Hägerstrand (1952, 1967) was the first scholar to
intensively analyze spatial diffusion pro cesses. Based on the empirical
observations of the regional spread of process innovations, he
hypothesized that new technology is first implemented in the `center' --
the large agglom erations -- and then moves `down' the spatial
hierarchy, the last stop being the `periphery', i.e.~remote and sparsely
populated regions. There are at least two factors that may explain this
pattern: first, the transfer of tacit knowledge via face-to-face contact
and second, the absorptive capacity of a region.

\citet{bento2015spatial} The concept of absorptive capacity has also
been applied in the TIS context in terms of the con-struction of
capacity needed to transfer and implement a new technology in the
receiving country (cf.Van Alphen, 2011), with repercussions for the
nature of system functions

\citet{bednarz2020pulled} larger local technological niches (at least
from a demand perspective) are more likely to be formed in urban areas
than in rural ones, which is in line with the work of Ha¨gerstrand
(1965a,b). However, size is not the only regional characteristic that
may matter in this context. Regions also differ in accumulated
experiences and the presence of tacit knowledge with respect to
products, as well as in actors' propensity share this knowledge among
producers (Martin et al., 2019)

@@

\citet{morrill2020spatial} Spatial diffusion is the process by which
behavior or characteristics of the landscape change as a result of what
happens elsewhere earlier. Spatial diffusion is the spread of the
phenomenon, over space and timed, from limited origins

@@

\citet{leibowicz2016representing} Diffusion is a spatial (Hagerstrand,
1967) as well as a temporal process, and historical evidence confirms
that technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies (Victor, 1993).

@@

\citet{grubler1990rise} Later Hagerstrand conceptualized physical
``barrier'' effects like lakes or uninhabited areas, which, in addition
to distance, act as further retarding effects on diffusion. These are
formalized in the form of ``zero'' or ``half'' contact multiplicators on
the (distance decaying) message flows.

\citet{grubler1990rise} With respect to the formalization of the
communication flows Hagerstrand defines a ``mean information field''
(MIF), in which the probability of communication is a negative function
of distance between individuals

\citet{perkins2005international} Theoretical models disagree as to why
firms adopt innovations at different times. Epidemic models emphasize
information (Griliches 1957). Certain firms are hypothesized to adopt
earlier because they come into contact with, and learn from, adopters of
the new technology before others. Economic models, on the other hand,
predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
Firms adopt technologies at different times because they differ with
respect to various organizational and environmental variables
influencing the economic returns from adoption (Blackman 1999).

@@

\citet{ding2010modeling} Rogers (1995) characterized early adopters as
knowledgeable risk takers. Griliches (1957) looked at the rate of return
for early adopters and characterized them as profit maximizers. The
geographer Hagerstrand (1967) was focusing on the character of early
adopters in the transmittal process that produces follow-on participants
in the process. Space was treated as a contiguity system with embedded
characteristics producing barrier elements to the diffusion process.

\citet{ding2010modeling} REVIEW ON Technological Diffusion

\citet{perkins2011internet} Two main mechanisms are identified in the
literature to explain diffusion: (1) epidemic-type dynamics whereby
contact with previous adopters stimulates uptake as po tential adopters
learn about a new innovation; and (2) economic-type mechanisms whereby
potential users adopt a new innovation as it be comesmore profitable,
useful, or valuable, with uptake characteristically spreading as costs
be come lower, performance improves, or the po tential uses of the
innovation grow over time.

@@

\citet{wilson201281} Logistic growth describes an initial period of
gradual diffusion as a technology is introduced as a new commercial
application, moving then through a rapid, exponential growth phase,
before slowing and eventually saturating \citep{grubler1999dynamics}.
The substitution of incumbent technologies by new competitors leads to
subsequent decline and eventual obsolescence.

@@

\citet{beardsell1999spatial} To characterize evolution of the computer
industry, we examine the distribution of relative employment across
cities in 1977 and how that distribution changes over time to 1992

@@

\citet{bednarz2020pulled} spatial innovation diffusion By using Bayesian
survival models with time-dependent data of wind turbine deployment and
firm foundation for 402 German regions between the years 1970 and 2015,
we show that the spatial evolution of the German wind energy industry
was more strongly influenced by local demand-- pull than local
supply--push processes.

@@

\citet{perkins2005international} analyze whether the rate at which new
producer technologies diffuse is significantly influenced by (1)
latecomer advantage and (2) engagement with the global economy via trade
and foreign investment

@@

\citet{fritsch2015new} analyze the spatial diffusion of laser technology
research in West Germany from 1960, when this technology began, until
2005.

\citet{perkins2011internet}

\citet{lengyel2020role} We see that the OSN spread almost exclusively
from the original location (the capital Budapest, with an order of
magnitude more inhabitants than the next size town) to various parts of
the country in the early phase of the life-cycle. Later, diffusion
became less mono-centric and other towns also emerged as spreaders. Our
findings support the idea that spreading initially happens to large
distances and becomes more local over time. This is illustrated and
discussed later in Figure 2I.

\citet{PAPAGIANNIDIS2015308}

\citet{leibowicz2016representing} for energy There is no generally
accepted theory that explains diffusion rate heterogeneity across
technologies, but several factors are considered important. Greater unit
scale and larger market size contribute to slower diffusion.
Requirements for interrelated technologies or complex infrastructures
also hinder the diffusion process (Grubler, 2012).

@@

\citet{leibowicz2016representing} Empirical evidence supports the
validity of Schmidt's Law over a wide range of technologies, time
periods, and geographical contexts. A recent meta-analysis of technology
up-scaling found that diffusion accelerated moving from the core to the
rim and periphery for technologies as diverse as natural gas power, oil
refineries, and automobiles (Wilson, 2009). One historical example that
conforms particularly well to Schmidt's Law is the diffusion of coal
power in Europe (Grubler, 2012). England emerged as the core region for
coal power because it had legal and economic institutions that
incentivized scientific pursuits, domestic coal reserves, and a clear
industrial motivation to replace water power with coal. \ldots{}

\citet{ding2010modeling} model the spatial diffusion of mobile
telecommunications in China as well as its determinants. regional
socioeconomic characteristics play an important role in determining the
timing, speed, and level of mobile telecommunications diffusion in China

\citet{bento2018time} explore What determines the duration of formative
phases for energy innovations in different markets? We are interested
both in initial markets (also: core, lead, first mover, early adopter)
where formative phases prepare technologies for mass commercialization,
and in follower markets (also: periphery, lag, late adopter) where
accelerated formative phases may benefit from diffusion and spillovers.

@@

\citet{wilson201281} growth function description in p.~86

@@

\textbf{minimum R2 = 95\%}, see also \citet{grubler1990rise}

@@

Moran's I as \citet{ding2010modeling} \textbf{TODO} for t\_0, diffusion
speed

@@

contrary to results from future studies regarding social media
\citep{lengyel2020role}, web technologies did not exclusively spread
from a central location.

\renewcommand\refname{References}
  \bibliography{bibliography}

>>>>

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{University of Bristol and The Alan Turing Institute}
}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS
journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Emmanouil Tranos\textsuperscript{\getAff{University of Bristol and The
Alan Turing Institute}}\textsuperscript{*}\\
\bigskip
\textbf{\getAff{University of Bristol and The Alan Turing
Institute}}University of Bristol, School of Geographical Sciences,
Bristol, UK; The Alan Turing Institute, London, UK\\
\bigskip
* Corresponding author: e.tranos@bristol.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
blah blah

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
blah blah

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{sec:1}{%
\section{Introduction}\label{sec:1}}

Geographers were always interested in how new technologies and
innovations diffuse across space and, importantly, how such
spatio-temporal processes can be modelled. The seminal contribution of
{[}1{]} is illustrative of this early interest. However, the torch of
exploring and modelling such processes had been passed to other
disciplines such as economics, business studies and sociology well
before the `cultural turn' of economic geography {[}2{]}. A potential
explanation of lack of studies exploring the diffusion of new and, more
specifically, digital technologies across \emph{both} space and time can
be attributed to the scarcity of relevant and granular enough data. As
{[}3{]} highlight, digital activities are hardly ever captured in
official data.

@@

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time use this information to make predictions
regarding the diffusion of related future technologies. As per {[}4{]} a
variety of actors have a direct interest in gaining such knowledge
including network equipment suppliers; network operators, regulatory and
local authorities. And despite the spatial heterogeneity of such
diffusion mechanisms and the policy relevance, there are very limited
attempts in the literature to analyse the diffusion of new digital
technologies at a granular level.

@@

{[}3{]} Because new digital activities are rarely---if ever---captured
in official state data, researchers must rely on information gathered
from alternative sources. With this in mind, the overarching aim of this
article is outlining an approach for analyzing the ``new spaces and
geographies'' of digital phenomena and practices.

{[}4{]} There are numerous agents who would benefit from accurate
predictions of introduction times. These include: suppliers of network
equipment for production planning; suppliers of mobile phones who need
to have inventories of appropriately configured handsets; international
network planners who need to provide sufficient bandwidth for
communications between different countries; network operators and
regulatory authorities.

@@

{[}5{]} Hägerstrand (1952, 1967) was the first scholar to intensively
analyze spatial diffusion pro cesses. Based on the empirical
observations of the regional spread of process innovations, he
hypothesized that new technology is first implemented in the `center' --
the large agglom erations -- and then moves `down' the spatial
hierarchy, the last stop being the `periphery', i.e.~remote and sparsely
populated regions. There are at least two factors that may explain this
pattern: first, the transfer of tacit knowledge via face-to-face contact
and second, the absorptive capacity of a region.

{[}6{]} The concept of absorptive capacity has also been applied in the
TIS context in terms of the con-struction of capacity needed to transfer
and implement a new technology in the receiving country (cf.Van Alphen,
2011), with repercussions for the nature of system functions

{[}7{]} larger local technological niches (at least from a demand
perspective) are more likely to be formed in urban areas than in rural
ones, which is in line with the work of Ha¨gerstrand (1965a,b). However,
size is not the only regional characteristic that may matter in this
context. Regions also differ in accumulated experiences and the presence
of tacit knowledge with respect to products, as well as in actors'
propensity share this knowledge among producers (Martin et al., 2019)

@@

{[}8{]} Spatial diffusion is the process by which behavior or
characteristics of the landscape change as a result of what happens
elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
over space and timed, from limited origins

@@

{[}9{]} Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
process, and historical evidence confirms that technologies diffuse at
different times, at different rates, and to different extents in
different places, and can be significantly influenced by policies
(Victor, 1993).

@@

{[}10{]} Later Hagerstrand conceptualized physical ``barrier'' effects
like lakes or uninhabited areas, which, in addition to distance, act as
further retarding effects on diffusion. These are formalized in the form
of ``zero'' or ``half'' contact multiplicators on the (distance
decaying) message flows.

{[}10{]} With respect to the formalization of the communication flows
Hagerstrand defines a ``mean information field'' (MIF), in which the
probability of communication is a negative function of distance between
individuals

{[}2{]} Theoretical models disagree as to why firms adopt innovations at
different times. Epidemic models emphasize information (Griliches 1957).
Certain firms are hypothesized to adopt earlier because they come into
contact with, and learn from, adopters of the new technology before
others. Economic models, on the other hand, predominantly emphasize firm
heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
different times because they differ with respect to various
organizational and environmental variables influencing the economic
returns from adoption (Blackman 1999).

@@

{[}11{]} Rogers (1995) characterized early adopters as knowledgeable
risk takers. Griliches (1957) looked at the rate of return for early
adopters and characterized them as profit maximizers. The geographer
Hagerstrand (1967) was focusing on the character of early adopters in
the transmittal process that produces follow-on participants in the
process. Space was treated as a contiguity system with embedded
characteristics producing barrier elements to the diffusion process.

{[}11{]} REVIEW ON Technological Diffusion

{[}12{]} Two main mechanisms are identified in the literature to explain
diffusion: (1) epidemic-type dynamics whereby contact with previous
adopters stimulates uptake as po tential adopters learn about a new
innovation; and (2) economic-type mechanisms whereby potential users
adopt a new innovation as it be comesmore profitable, useful, or
valuable, with uptake characteristically spreading as costs be come
lower, performance improves, or the po tential uses of the innovation
grow over time.

@@

{[}13{]} Logistic growth describes an initial period of gradual
diffusion as a technology is introduced as a new commercial application,
moving then through a rapid, exponential growth phase, before slowing
and eventually saturating {[}14{]}. The substitution of incumbent
technologies by new competitors leads to subsequent decline and eventual
obsolescence.

@@

{[}15{]} To characterize evolution of the computer industry, we examine
the distribution of relative employment across cities in 1977 and how
that distribution changes over time to 1992

@@

{[}7{]} spatial innovation diffusion By using Bayesian survival models
with time-dependent data of wind turbine deployment and firm foundation
for 402 German regions between the years 1970 and 2015, we show that the
spatial evolution of the German wind energy industry was more strongly
influenced by local demand-- pull than local supply--push processes.

@@

{[}2{]} analyze whether the rate at which new producer technologies
diffuse is significantly influenced by (1) latecomer advantage and (2)
engagement with the global economy via trade and foreign investment

@@

{[}5{]} analyze the spatial diffusion of laser technology research in
West Germany from 1960, when this technology began, until 2005.

{[}12{]}

{[}16{]} We see that the OSN spread almost exclusively from the original
location (the capital Budapest, with an order of magnitude more
inhabitants than the next size town) to various parts of the country in
the early phase of the life-cycle. Later, diffusion became less
mono-centric and other towns also emerged as spreaders. Our findings
support the idea that spreading initially happens to large distances and
becomes more local over time. This is illustrated and discussed later in
Figure 2I.

{[}17{]}

{[}9{]} for energy There is no generally accepted theory that explains
diffusion rate heterogeneity across technologies, but several factors
are considered important. Greater unit scale and larger market size
contribute to slower diffusion. Requirements for interrelated
technologies or complex infrastructures also hinder the diffusion
process (Grubler, 2012).

@@

{[}9{]} Empirical evidence supports the validity of Schmidt's Law over a
wide range of technologies, time periods, and geographical contexts. A
recent meta-analysis of technology up-scaling found that diffusion
accelerated moving from the core to the rim and periphery for
technologies as diverse as natural gas power, oil refineries, and
automobiles (Wilson, 2009). One historical example that conforms
particularly well to Schmidt's Law is the diffusion of coal power in
Europe (Grubler, 2012). England emerged as the core region for coal
power because it had legal and economic institutions that incentivized
scientific pursuits, domestic coal reserves, and a clear industrial
motivation to replace water power with coal. \ldots{}

{[}11{]} model the spatial diffusion of mobile telecommunications in
China as well as its determinants. regional socioeconomic
characteristics play an important role in determining the timing, speed,
and level of mobile telecommunications diffusion in China

{[}18{]} explore What determines the duration of formative phases for
energy innovations in different markets? We are interested both in
initial markets (also: core, lead, first mover, early adopter) where
formative phases prepare technologies for mass commercialization, and in
follower markets (also: periphery, lag, late adopter) where accelerated
formative phases may benefit from diffusion and spillovers.

@@

{[}13{]} growth function description in p.~86

@@

\textbf{minimum R2 = 95\%}, see also {[}10{]}

@@

Moran's I as {[}11{]} \textbf{TODO} for t\_0, diffusion speed

@@

contrary to results from future studies regarding social media {[}16{]},
web technologies did not exclusively spread from a central location.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-hagerstrand1968innovation}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{Hagerstrand T et al. Innovation diffusion as a spatial
process. Innovation diffusion as a spatial process. Chicago, USA: Univ.
Chicago Press.; 1968; }

\leavevmode\vadjust pre{\hypertarget{ref-perkins2005international}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Perkins R, Neumayer E. The international diffusion of
new technologies: A multitechnology analysis of latecomer advantage and
global economic integration. Annals of the Association of American
Geographers. Taylor \& Francis; 2005;95: 789--808. }

\leavevmode\vadjust pre{\hypertarget{ref-zook2022mapping}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Zook M, McCanless M. Mapping the uneven geographies of
digital phenomena: The case of blockchain. The Canadian Geographer/Le
G{é}ographe canadien. Wiley Online Library; 2022;66: 23--36. }

\leavevmode\vadjust pre{\hypertarget{ref-meade2021modelling}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Meade N, Islam T. Modelling and forecasting national
introduction times for successive generations of mobile telephony.
Telecommunications Policy. Elsevier; 2021;45: 102088. }

\leavevmode\vadjust pre{\hypertarget{ref-fritsch2015new}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Fritsch M, Medrano Echalar LF. New technology in the
region--agglomeration and absorptive capacity effects on laser
technology research in west germany, 1960--2005. Economics of Innovation
and New Technology. Taylor \& Francis; 2015;24: 65--94. }

\leavevmode\vadjust pre{\hypertarget{ref-bento2015spatial}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Bento N, Fontes M. Spatial diffusion and the formation
of a technological innovation system in the receiving country: The case
of wind energy in portugal. Environmental Innovation and Societal
Transitions. Elsevier; 2015;15: 158--179. }

\leavevmode\vadjust pre{\hypertarget{ref-bednarz2020pulled}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Bednarz M, Broekel T. Pulled or pushed? The spatial
diffusion of wind energy between local demand and supply. Industrial and
Corporate Change. Oxford University Press; 2020;29: 893--916. }

\leavevmode\vadjust pre{\hypertarget{ref-morrill2020spatial}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Morrill R, Gaile GL, Thrall GI. Spatial diffusion. Sage
Publications; 1988; }
